---
title: "MA615/415 Midterm Project"
subtitle: "County-level oil and gas production"
author: "Chenlu Liu"
output: pdf_document
header-includes: \usepackage{float}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Set Up: Packages and Functions

We will need to use the following packages: `"stringr"`, `"ggplot2"`, and
`"dplyr"`. install them first (outside this Rmd).

install.packages(`"stringr"`)

install.packages(`"ggplot2"`)

install.packages(`"dplyr"`)

```{r}
# use this code chunk to load all the packages that you will be using
library(dplyr)
library(plyr)
library(ggplot2)
# (do not include commands to install the packages)
```



# Import Raw Data in R

I already downloaded a copy of the CSV files with the
raw data : 

- `oilgascounty.csv`


Now, use one of the reading table functions to import the data sets in R. 
Do NOT convert strings into factors.

```{r}


# read in oilgascounty.csv
oilgascounty = read.csv("D:\\midterm\\oilgascounty.csv", stringsAsFactors = FALSE)

# check structure with str()

str(oilgascounty)

```
__FIPS__:  Five digit Federal Information Processing Standard (FIPS) code (numeric)

__geoid__:  FIPS code with leading zero (string)

__Stabr__:	State abbreviation (string)

__County_Name__:  County name (string)

__Rural_Urban_Continuum_Code_2013__:	Rural-urban Continuum Code 2013 (see code descriptions)

__Urban_Influence_2013__:	Urban Influence Code, 2013 (see code descriptions)

__Metro_Nonmetro_2013__:	Metro-nonmetro 2013 (0=nonmetro, 1=metro)

__Metro_Micro_Noncore_2013__:	Metro-Micro-Noncore Indicator 2013 (0=nonmetro noncore, 1=nonmetro micropolitan, 2=metropolitan)

__oil2000__, __oil2001__, ..., __oil2011__:	Annual gross withdrawals (barrels) of crude oil, for the year specified in the variable name

__gas2000__, __gas2001__, ..., __gas2011__:	Annual gross withdrawals (thousand cubic feet) of natural gas, for the year specified in the variable name

__oil_change_group__:	Categorical variable based upon change in the dollar value of oil production, 2000-11. Values are H_Growth (>=$20 million),H_ Decline (<=-$20 million), Status Quo (change between +/- $20 million)

__gas_change_group__:	Categorical variable based upon change in the dollar value of natural gas production, 2000-11. Values are H_Growth (>=$20 million), H_Decline (<=-$20 million), Status Quo (change between +/- $20 million)

__oil_gas_change_group__:	Categorical variable based on the change in the dollar value of the sum of oil and natural gas production, 2000-11. Values are H_Growth (>=$20 million),H_ Decline (<=-$20 million), Status Quo (change between +/- $20 million)

# Cleaning raw data of County-level oil and gas production
First, we  summary the data 
```{r}
summary(oilgascounty)
```
then we
count
the number of __oil_gas_change_group__ in each  (__Metro_Nonmetro_2013__) with the tidy __count__ function.
```{r}
count(oilgascounty, c("Metro_Nonmetro_2013", "oil_gas_change_group"))


```
count
the number of __oil_gas_change_group__ in each  (__Metro_Micro_Noncore_2013__) with the tidy __count__ function.
```{r}

count(oilgascounty, c("Metro_Micro_Noncore_2013", "oil_gas_change_group"))


```
Then we remove missing (and hence uninformative for our purpose) values with subset.
```{r}
oil_group2 <- subset(oilgascounty,  oil2000>0&oil2001>0&oil2002>0&oil2003>0&oil2004>0&oil2005>0
                                &oil2006>0&oil2007>0&oil2008>0&oil2009>0&oil2010>0&oil2011>0)

oil_group3=apply(oil_group2[,9:20],2,sum)
new_data=data.frame(time=2000:2011,stype=rep("oil",12),prudction=as.numeric(oil_group3))
gas_group2 <- subset(oilgascounty,  gas2000>0&gas2001>0&gas2002>0&gas2003>0&gas2004>0&gas2005>0
                                &gas2006>0&gas2007>0&gas2008>0&gas2009>0&gas2010>0&gas2011>0)
gas_group3=apply(gas_group2[,9:20],2,sum)
new_data=rbind(new_data,data.frame(time=2000:2011,stype=rep("gas",12),prudction=as.numeric(gas_group3)))


```

# Data Visualization

Use the `records` data frame, and functions in `"ggplot2"`, to create charts 
similar to those displayed in the PDF with the instructions for this project:

```{r}
qplot(time,prudction,  data = new_data, geom=c("line","point"),
      main="oil and gas Tidy Data ",colour = stype)
      
  ```    
      